% https://github.com/open-mpi/ompi/blob/master/ompi/mca/pml/pml.h 
 * P2P Management Layer (PML)
 *
 * An MCA component type that provides the P2P interface functionality
 * required by the MPI layer. The PML is a relatively thin layer that
 * primarily provides for the fragmentation and scheduling of messages
 * over multiple transports (instances of the Byte Transfer Layer
 * (BTL) MCA component type) as depicted below:
 *
 *   ------------------------------------
 *   |                MPI               |
 *   ------------------------------------
 *   |                PML               |
 *   ------------------------------------
 *   | BTL (TCP) | BTL (SM) | BTL (...) |
 *   ------------------------------------
 *
 * A single PML component is selected by the MCA framework during
 * library initialization. Initially, all available PMLs are loaded
 * (potentially as shared libraries) and their component open and init
 * functions called.  The MCA framework selects the component
 * returning the highest priority and closes/unloads any other PML
 * components that may have been opened.
 *
 * After all of the MCA components are initialized, the MPI/RTE will
 * make downcalls into the PML to provide the initial list of
 * processes (ompi_proc_t instances), and notification of changes
 * (add/delete).
 *
 * The PML module must select the set of BTL components that are to be
 * used to reach a given destination. These should be cached on a PML
 * specific data structure that is hung off the ompi_proc_t.
 *
 * The PML should then apply a scheduling algorithm (round-robin,
 * weighted distribution, etc), to schedule the delivery of messages
 * over the available BTLs.